---
title: "OISST CMIP6 Bias Correction"
author: "Adam Kemberling + Matt Dzaugis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

# Set knitr options
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, comment = NA)
options(knitr.kable.NA = '')

###__ Packages  ####
library(here)
library(raster)
library(janitor)
library(gmRi)
library(patchwork)
library(tidyverse)
library(knitr)



# Load the build code and stratification function
box_paths  <- research_access_paths(os.use = "unix")
mills_path <- box_paths$mills
res_path   <- box_paths$res
oisst_path <- paste0(res_path, "OISST/oisst_mainstays/")

#### Set theme  ####
theme_set(theme_minimal())

#  color palette for quick raster displays
temp_pal <- rev(RColorBrewer::brewer.pal(n = 10, name = "RdBu"))

```

`r use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")`

# Bias Correcting CMIP6 SST with OISST 

This doc is a sandbox for testing approaches for pairing OISST climate means with CMIP6 SST data for bias corrections.

### Loading Data Resources  

```{r}

# Load the climatology for 1982-2011
oisst_clim_82 <- stack(str_c(oisst_path, "daily_climatologies/daily_clims_1982to2011.nc"))

# Load a CMIP6 file for testing:
cmip_file <- paste0(res_path, "CMIP6/TestFiles/tos_Omon_CanESM5_historical_r1i1p2f1_gn_195501-201412.nc.1x1.nc")
cmip_test <- stack(cmip_file)
#plot(cmip_test$X1955.01.16, col = temp_pal, main = "CMIP6 1955-01-16")

```

# Prepare data for bias correction


## OISST Prep

 - Remove unwanted areas to speed things up
 - Convert daily climate means to monthly averages

```{r}


# Crop Area to CMIP Area
# Northeast Area
ne_area <- extent(c(260, 320, 20, 70))
clim_cropped <- crop(oisst_clim_82, ne_area)
rm(oisst_clim_82)



# Use number key to match months to corresponding the day of year,
# day of year in climatology honors the 60th as only feb 29th
# in this system march 1st is always 61, and Dec 31st is always 366
months <- list(
  "Jan" = c(1:31),
  "Feb" = c(32:60),
  "Mar" = c(61:91),
  "Apr" = c(92:121),
  "May" = c(122:152),
  "Jun" = c(153:182),
  "Jul" = c(183:213),
  "Aug" = c(214:244),
  "Sep" = c(245:274),
  "Oct" = c(275:305),
  "Nov" = c(306:335),
  "Dec" = c(336:366))
month_abbrevs <- names(months)


# Put a capital X in front of each of them for stack names
month_key <- map(months, ~ str_c("X", .x))


# Use month ket to subset the stack for each month
# get the mean, then restack
oisst_month_avgs <- map(month_key, function(month_indices){
  month_mean <- mean(clim_cropped[[month_indices]])}) %>% 
  stack() %>% 
  setNames(month_abbrevs)


# Plot a month
plot(oisst_month_avgs$Jun, main = "June 1982-2011 Climate Avg. OISST", col = temp_pal)



```


## CMIP Prep {.tabset .tabset-pills}

 - Remove unwanted areas to speed things up   
 - Get monthly climate averages to get anomalies   
 - Compare monthly cmip climate to oisst climate to get bias   
 - Use bias to adjust anomalies   
 
### Crop CMIP

```{r}

# Crop CMIP
cmip_cropped <- crop(cmip_test, ne_area)
rm(cmip_test)

# # plot check
plot(cmip_cropped$X1955.01.16, main = "CMIP6 1955-01-16", col = temp_pal)
```

### CMIP Climatology

```{r}
####  Make Monthly Climatology  ####
clim_years <- as.character(c(1982:2011))
cmip_layers <- names(cmip_cropped)

# Pull layers of cmip data for the years of interest for climatology
cmip_clim_years <- cmip_cropped[[which(str_sub(cmip_layers, 2,5) %in% clim_years)]]
clim_year_layers <- names(cmip_clim_years)

# Get strings to match the months, set names as abbreviations for later
month_labels <- str_pad(c(1:12), width = 2, pad = "0", "left")
month_labels <- setNames(month_labels, month_abbrevs)

# Loop through the months, getting mean across the climatology period
cmip_clim <- map(month_labels, function(month_index){
  
  # Indices for the month
  cmip_month_indices <- which(str_sub(clim_year_layers, 7,8) == month_index)
  
  # Mean across years
  monthly_clim <- mean(cmip_clim_years[[cmip_month_indices]])
  return(monthly_clim)
}) %>% stack()


# Plot CMIP Monthly Climate
plot(cmip_clim$Jun, main = "June 1982-2011 CMIP Climate Avg. SST", col = temp_pal)
```

### CMIP Anomalies

```{r}
####  CMIP Anomalies  ####

# just need to loop through months again to pull
cmip_anoms <- imap(month_labels, function(month_index, month_abbrev){
  
  # Use it to get month from climatology
  clim_month_data <- cmip_clim[[month_abbrev]]
  
  # Get indices for CMIP layers that match the month
  cmip_month_indices <- which(str_sub(cmip_layers, 7, 8) == month_index)
  cmip_month_labels  <- cmip_layers[cmip_month_indices]
  
  # Pull all layers from the cropped cmip stack with that month
  month_layers <- cmip_cropped[[cmip_month_indices]]
  
  # Subtract climate average to get anomalies
  month_anoms <- month_layers - clim_month_data
  month_anoms <- setNames(month_anoms, cmip_month_labels)
  
  
}) %>% stack()


# Plot Test
plot(cmip_anoms$X1955.01.16, main = "CMIP6 Anomalies 1955-01-16", col = temp_pal)

```



## Bias Correction testing

Do a quick one off of for January of cmip vs january of oisst. This should highlight the difference (if any) of doing the delta method vs. just removing the difference between climates directly.


```{r}
# Get OISST Climatology for the whole region
oisst_clim_timeline <- cellStats(oisst_month_avgs, mean)
oisst_clim_timeline <- as.data.frame(oisst_clim_timeline) %>% rownames_to_column(var = "month")

# Same deal for cmip climatology
cmip_clim_timeline <- cellStats(cmip_clim, mean)
cmip_clim_timeline <- as.data.frame(cmip_clim_timeline) %>% rownames_to_column(var = "month")

# Pull one year of CMIP
test_year <- "2000"
which_cmip <- which(str_sub(names(cmip_cropped), 2, 5 ) == test_year) 
cmip_2020_timeline <- cellStats(cmip_cropped[[which_cmip]], mean, na.rm = T)
cmip_2020_timeline <- as.data.frame(cmip_2020_timeline) 
cmip_2020_timeline$month <- month_abbrevs
  
# # Pull the anomalies that year as well as a validation of that
# which_cmip_anoms <- which(str_sub(names(cmip_anoms), 2, 5 ) == test_year) 
# cmip_2020_anom_timeline <- cellStats(cmip_anoms[[which_cmip_anoms]], mean, na.rm = T)
# cmip_2020_anom_timeline <- as.data.frame(cmip_2020_anom_timeline)
# cmip_2020_anom_timeline$month <- month_abbrevs


# Plot the bias
all_timelines <- left_join(oisst_clim_timeline, cmip_clim_timeline, by = "month") %>% 
  left_join(cmip_2020_timeline, by = "month") %>% 
  #left_join(cmip_2020_anom_timeline, by = "month") %>% 
  pivot_longer(names_to = "data_source", values_to = "temp", cols = 2:4) %>% 
  mutate(month = factor(month, levels = month_abbrevs))

#
ggplot(all_timelines, aes(month, temp, color  = data_source)) +
  geom_line(aes(group = data_source)) +
  labs(x = "", y = "Temperature")


```

