---
title: "OISST CMIP6 Bias Correction"
author: "Adam Kemberling + Matt Dzaugis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: show
editor_options: 
  chunk_output_type: console

---

```{r setup, include=FALSE}

# Set knitr options
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, comment = NA)
options(knitr.kable.NA = '')

###__ Packages  ####
library(here)
library(raster)
library(janitor)
library(gmRi)
library(patchwork)
library(tidyverse)
library(knitr)



# Load the build code and stratification function
box_paths  <- research_access_paths(os.use = "unix")
mills_path <- box_paths$mills
res_path   <- box_paths$res
oisst_path <- paste0(res_path, "OISST/oisst_mainstays/")

#### Set theme  ####
theme_set(theme_minimal())

#  color palette for quick raster displays
temp_pal <- rev(RColorBrewer::brewer.pal(n = 10, name = "RdBu"))

####  Functions  ####
source(here("CMIP6_processing/R/sdm_workflow_funs.R"))

```

`r use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")`

# Bias Correcting CMIP6 SST with OISST 

This doc is a sandbox for testing approaches for pairing OISST climate means with CMIP6 SST data for bias corrections.

### Loading Data Resources  

```{r}

# Load the climatology for 1982-2011
oisst_clim_82 <- import_oisst_clim(climatology_period = "1982-2011")
oisst_clim_91 <- import_oisst_clim(climatology_period = "1991-2020")


# Load a CMIP6 file for testing:
cmip_cropped <- import_cmip_sst(cmip_file = "tester") 

# These have been cropped by default from the import functions

# Plot a tester
#plot(cmip_test$X1955.01.16, col = temp_pal, main = "CMIP6 1955-01-16")

```

# Prepare data for bias correction


## OISST Prep

 - Remove unwanted areas to speed things up
 - Convert daily climate means to monthly averages

```{r}


# Use months_from_clim() to get the monthly averages

# Default key uses modified ordinal days
oisst_month_avgs    <- months_from_clim(clim_source = oisst_clim_82, month_layer_key = NULL)
oisst_month_avgs_91 <- months_from_clim(clim_source = oisst_clim_91) 


# Plot a month from each
#par(mfrow = c(2,1))
plot(oisst_month_avgs$Jun, main = "June 1982-2011 Climate Avg. OISST", 
     col = temp_pal, xlim = c(260, 320))
plot(oisst_month_avgs_91$Jun, main = "June 1991-2020 Climate Avg. OISST", col = temp_pal)



```


## CMIP Prep {.tabset .tabset-pills}

 - Remove unwanted areas to speed things up   
 - Get monthly climate averages to get anomalies   
 - Compare monthly cmip climate to oisst climate to get bias   
 - Use bias to adjust anomalies   
 
### Crop CMIP

```{r}
# # plot check
plot(cmip_cropped$X1955.01.16, main = "CMIP6 1955-01-16", col = temp_pal)
```

### CMIP Climatology

```{r}
####  Make Monthly Climatology  ####
cmip_clim <- cmip_to_clim(cmip = cmip_cropped)


# Plot CMIP Monthly Climate
plot(cmip_clim$Jun, main = "June 1991-2020\nCMIP Climate Avg. SST", 
     col = temp_pal)
```

### CMIP Anomalies

```{r}
# Match months to climatology to get anomalies
cmip_anoms <- cmip_get_anomalies(cmip_data = cmip_cropped, cmip_clim = cmip_clim)


# Plot Test
plot(cmip_anoms$X1955.01.16, main = "CMIP6 Anomalies 1955-01-16", col = temp_pal)

```



## Whole-Area Bias Testing

Do a quick one off of for January of cmip vs january of oisst. This should highlight the difference (if any) of doing the delta method vs. just removing the difference between climates directly.


```{r}
# Get OISST Climatology for the whole region

####  Switching to 1991-2020 ####
oisst_clim_timeline <- cellStats(oisst_month_avgs_91, mean)
oisst_clim_timeline <- as.data.frame(oisst_clim_timeline) %>% rownames_to_column(var = "month")

# Same deal for cmip climatology
cmip_clim_timeline <- cellStats(cmip_clim, mean)
cmip_clim_timeline <- as.data.frame(cmip_clim_timeline) %>% rownames_to_column(var = "month")

# Pull one year of CMIP
test_year <- "2000"
which_cmip <- which(str_sub(names(cmip_cropped), 2, 5 ) == test_year) 
cmip_2020_timeline <- cellStats(cmip_cropped[[which_cmip]], mean, na.rm = T)
cmip_2020_timeline <- as.data.frame(cmip_2020_timeline) 
cmip_2020_timeline$month <- month_abbrevs
  
# # Pull the anomalies that year as well as a validation of that
# which_cmip_anoms <- which(str_sub(names(cmip_anoms), 2, 5 ) == test_year) 
# cmip_2020_anom_timeline <- cellStats(cmip_anoms[[which_cmip_anoms]], mean, na.rm = T)
# cmip_2020_anom_timeline <- as.data.frame(cmip_2020_anom_timeline)
# cmip_2020_anom_timeline$month <- month_abbrevs


# Plot the bias
all_timelines <- left_join(oisst_clim_timeline, cmip_clim_timeline, by = "month") %>% 
  left_join(cmip_2020_timeline, by = "month") %>% 
  #left_join(cmip_2020_anom_timeline, by = "month") %>% 
  pivot_longer(names_to = "data_source", values_to = "temp", cols = 2:4) %>% 
  mutate(month = factor(month, levels = month_abbrevs))

#
ggplot(all_timelines, aes(month, temp, color  = data_source)) +
  geom_line(aes(group = data_source)) +
  geom_point(size = 1) +
  scale_color_gmri() +
  labs(x = "", 
       y = "Temperature", 
       caption = "Whole-area difference in climatology.")


```


## Bias correcting work flows


### The delta method


This is what we have done in the past, with guidance from Mike and Andy. The process involves three steps, two of which are already done:
  1. Calculating climate model climatology
  2. Calculate climate model deltas from the climate model climatology
  3. Apply the deltas of the climate model to the observed OISST climatology for the same reference period


```{r}

# The file that has the anomalies from the climate model 
# (climate model projected temps - climate model climatology) -- 
# these are also "deltas"

# cmip_anoms


# Grabbing how I had done this before requires a litle adjusting.
# Resampling cmip anomalies to match oisst grid
cmip_anoms_regridded <- resample_grid(starting_grid = cmip_anoms, 
                               desired_grid = oisst_month_avgs_91, 
                               method = "bilinear")



# Use delta-method to bias correct the data
cmip_delta_bias_corrected <- delta_method_bias_correct(cmip_grid = cmip_anoms_regridded, 
                                                       reference_climatology = oisst_month_avgs_91)





# Get difference from the "biased" original?
# Resampling to match resolutions
cmip_cropped_ds <- resample(cmip_cropped, oisst_month_avgs_91[[1]], method = "bilinear")


# Now one example difference
cmip_diffs <- cmip_cropped_ds$X2014.04.16 - cmip_delta_bias_corrected$X2014.04.16
par(mfrow = c(1, 3))
plot(cmip_cropped$X2014.04.16, 
     main = "CMIP6 2014-01-16\nOriginal", col = temp_pal, zlim = c(-1, 28)) 
plot(cmip_delta_bias_corrected$X2014.04.16, 
     main = "CMIP6 2014-01-16\nBias-corrected delta method", col = temp_pal, zlim = c(-1, 28))
plot(cmip_diffs, 
     main = "CMIP6 2014-01-16\nDifferences", col = temp_pal, zlim = c(-10, 10)) 

#dev.off()
```



### The "nudging" method?

An alternative way of thinking about this is to calculate a bias over the same reference period between the climate model temperatures and the observed temperatures. This bias would then be used to adjust the climate model projected temperatures.

```{r, eval = FALSE}

# First, get the two climatology stacks and make sure they have the same resolution. First, climatology based on the climate model projected temps and then OISST. Both calculated over 1982-2011 period.
cmip_clim_ds <- resample(cmip_clim, oisst_month_avgs_91[[1]], method = "bilinear") 
oisst_month_avgs_91


# Take the difference to get the "bias" during this baseline period
cmip_clim_bias <- cmip_clim_ds - oisst_month_avgs_91


# Now, adjust all of the climate model projected temperatures according to this bias. This is similar to applying anomalies to corresponding matching OISST month.


# Work on names
names(cmip_clim_bias) <- months_numeric
cmip_dates <- as.Date(gsub("[.]", "-", gsub("X", "", names(cmip_cropped_ds))))
cmip_cropped_z <- setZ(cmip_cropped_ds, cmip_dates)


# Alright, for every climate model projected temperature, adjust based on bias (difference between climate model and observed)
cmip_proj_out_alt <- lapply(seq(1:nlayers(cmip_cropped_z)), function(x) {
  cmip_layer      <- cmip_cropped_z[[x]]
  layer_match     <- match(format(getZ(cmip_cropped_z)[x], "%m"), gsub("X", "", 
                                                                       
                                                                       names(cmip_clim_bias)))
  cmip_bias_layer <- cmip_clim_bias[[layer_match]]
  delta_out       <- cmip_layer - cmip_bias_layer}) %>% 
  stack()
names(cmip_proj_out_alt) <- cmip_dates


# Get difference from the "biased" original?
cmip_diffs_alt<- cmip_cropped_ds$X2014.04.16 - cmip_proj_out_alt$X2014.04.16
par(mfrow = c(1, 3))
plot(cmip_cropped$X2014.04.16, main = "CMIP6 2014-01-16", col = temp_pal, zlim = c(-1, 28))
plot(cmip_proj_out_alt$X2014.04.16, main = "CMIP6 2014-01-16 nudged", col = temp_pal, zlim = c(-1, 28))
plot(cmip_diffs_alt, main = "CMIP6 2014-01-16 difference", col = temp_pal, zlim = c(-10, 10))


# Any difference??
par(mfrow = c(1,1))
cmip_appr_diff<- cmip_proj_out$X2014.04.16 - cmip_proj_out_alt$X2014.04.16
plot(cmip_appr_diff, main = "CMIP6 2014-01-16 Approach differences", col = temp_pal)

```




