---
title: "Bias-Corrected Regional CMIP Projections"
author: "Adam A. Kemberling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: hide
editor_options: 
  chunk_output_type: console

---

```{r setup, include=FALSE}
# Set knitr options
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, comment = NA)
options(knitr.kable.NA = '')

###__ Packages  ####
library(here)
library(ncdf4)
library(RNetCDF)
library(rnaturalearth)
library(raster)
library(janitor)
library(gmRi)
library(patchwork)
library(tidyverse)
library(knitr)
library(sf)
library(ggforce)



# Load the build code and stratification function
box_paths  <- research_access_paths()
res_path   <- box_paths$res
oisst_path <- box_paths$oisst_mainstays
cmip_path  <- box_path("RES_Data", "CMIP6")
soda_path  <- box_path("RES_Data", "SODA")

# Set theme  
theme_set(theme_bw())

#  color palette for quick raster displays
temp_pal <- rev(RColorBrewer::brewer.pal(n = 10, name = "RdBu"))

# Coastlines
world       <- ne_countries(scale = "medium", returnclass = "sf")
new_england <- ne_states("united states of america", returnclass = "sf")
canada      <- ne_states("canada", returnclass = "sf")
greenland   <- ne_states(country = "greenland", returnclass = "sf")

####  Functions  ####
source(here("CMIP6_processing/R/sdm_workflow_funs.R"))
```



`r use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")`


# Validating Bias-Correction of CMIP6 Data

This report exists to visually compare the bias-corrected CMIP6 data against the real-world observation data sources used to bias-correct them.

Bias corrected timelines for the sea surface temperature, sea surface salinity, bottom temperature, and bottom salinity will be compared against the reference observations (OISSTv2, SODA) to check how accurately they represent observed conditions, and whether that changes between our areas of interest. 


## Load Data {.tabset .tabset-pills}

Both the bias-corrections and the observational datasets (OISSTv2, SODA) have been prepared separately and are loaded below.


###  Target Areas {.tabset }

For clarity on what data is included/attributed to each region, the areas we focused on have been plotted below.

```{r get polygons}

# Getting Timeseries and Shapefile Locations via gmRi
# returns path to oisst timeseries & path to shapefile:
region_groups <- c("nelme_regions", "gmri_sst_focal_areas", "lme", "nmfs_trawl_regions")
region_lookup <- map(region_groups, function(region_group){
    mask_details <- get_timeseries_paths(region_group) }) %>% 
  setNames(region_groups)

# Grab a couple to use from each as contenders


# NMFS Trawl Regions
trawl_gb  <- read_sf(region_lookup$nmfs_trawl_regions$georges_bank$shape_path) 
trawl_gom <- read_sf(region_lookup$nmfs_trawl_regions$gulf_of_maine$shape_path)
trawl_mab <- read_sf(region_lookup$nmfs_trawl_regions$mid_atlantic_bight$shape_path)
trawl_sne <- read_sf(region_lookup$nmfs_trawl_regions$southern_new_england$shape_path)


# Load all the strata and just filter out the crap ones
trawl_full <- read_sf(str_c(res_path, "Shapefiles/BottomTrawlStrata/BTS_Strata.shp"))  %>% 
  clean_names() %>% 
  filter(strata >= 01010 ,
         strata <= 01760,
         strata != 1310,
         strata != 1320,
         strata != 1330,
         strata != 1350,
         strata != 1410,
         strata != 1420,
         strata != 1490) 


# DFO Data
dfo_path <- shared.path(group = "Mills Lab", folder = "Projects/DFO_survey_data/strata_shapefiles")
dfo_area <- read_sf(str_c(dfo_path, "MaritimesRegionEcosystemAssessmentBoundary.shp"))


# set overall zoom for maps
xlimz <- c(-76, -57)
ylimz <- c(35, 48)

# base map
base_map <- ggplot() +
  geom_sf(data = new_england) +
  geom_sf(data = canada) +
  geom_sf(data = greenland) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title.position = "top", title.hjust = 0.5)) +
  labs(color = "", 
       fill = "")
```

#### NMFS Trawl Areas

```{r}
# Build off of base map
base_map +
  geom_sf(data = st_union(trawl_full), aes(fill = "Trawl Survey Area"), alpha = 0.2) +
  geom_sf(data = st_union(trawl_gom), aes(fill = "Gulf of Maine Survey Area"), alpha = 0.2) +
  coord_sf(xlim = xlimz, ylim = ylimz) +
  labs(title = "NMFS Trawl Region")

```

####  DFO Survey Areas

```{r}

# Map dfo data 
base_map + 
  geom_sf(data = dfo_area, aes(fill = "DFO Ecosystems Assesment Boundary"), alpha = 0.2) +
  coord_sf(xlim = xlimz, ylim = ylimz) +
  labs(title = "DFO Region")

```

#### All Areas


```{r}
base_map +
  geom_sf(data = st_union(trawl_full), aes(fill = "Trawl Survey Area"), alpha = 0.2) +
  geom_sf(data = st_union(trawl_gom), aes(fill = "Gulf of Maine Survey Area"), alpha = 0.2) +
  geom_sf(data = dfo_area, aes(fill = "DFO Ecosystems Assesment Boundary"), alpha = 0.2) +
  coord_sf(xlim = xlimz, ylim = ylimz) +
  labs(title = "All Areas")
```


### SODA

```{r}
# SODA vars
soda_ssal  <- stack(str_c(soda_path, "SODA_Salt_Red.nc"))
soda_bsal  <- stack(str_c(soda_path, "SODA_Salt_Red_bottomLayer.nc"))
soda_btemp <- stack(str_c(soda_path, "SODA_Temp_Red_bottomLayer.nc"))

```



### OISSTv2

```{r}

# OISST 
data_window <- data.frame(lon = c(-72, -61),
                          lat = c(39, 46.1),
                          time = as.Date(c("1981-09-01", "2020-12-31")))

# Load Data
oisst <- oisst_window_load(data_window = data_window, 
                           anomalies = FALSE)


# Convert to Monthly
oisst <- stack(oisst)

# build a key for months using the layer names
month_key <- seq.Date(min(data_window$time), max(data_window$time), by = "month")
month_key <- str_c("X", month_key) %>% str_replace_all("-", ".")
month_key <- str_sub(month_key, 1, -4)

# Crop them
oisst_monthly <- map(month_key, function(wut_month){
  
  # Identify corresponding days in month
  which_layers  <- str_detect(names(oisst), wut_month)
  layer_indices <- which(which_layers)
  
  if(length(layer_indices) == 0) { print(str_c(length(layer_indices), " days for ", wut_month))  }
  
  # Get mean
  month_mean <- calc(oisst[[layer_indices]], fun = mean, na.rm = T)

  # set month as name
  month_mean <- setNames(month_mean, wut_month)
  return(month_mean)
  
  
}) 


# Stack
oisst_monthly <- stack(oisst_monthly)

# clean up
rm(oisst)
```


**Cropping Functions**

```{r}
####  Processing Functions
# Masking function for an area
mask_shape <- function(in_ras, in_mask){
  r1 <- crop(x = in_ras, y = in_mask)
  r2 <- mask(x = r1, mask = in_mask)
  return(r2)}

# Function to turn it into a dataframe using cellstats
timeseries_from_mask <- function(ras_in, var_name){
  ts <- cellStats(ras_in, mean, na.rm = T)
  ts <- ts %>% 
    as.data.frame() %>% 
    rownames_to_column() %>% 
    setNames(c("date", var_name)) %>% 
    mutate(date = str_remove(date, "X"),
           date = str_replace_all(date, "[.]", "-"))
  
  date_len <- str_length(ts$date[1])
  if(date_len == 7){
    ts <- ts %>% 
      mutate(date = str_c(date, "-15"),
             date = as.Date(date))
  } else{
    ts <- mutate(ts, date = as.Date(date))
  }
  
  return(ts)
}
```

**Processing Observed Variables**

```{r}
# Put observed variables in a list
observed_vars <- list(
  bot_temp  = soda_btemp,
  bot_sal   = soda_bsal,
  surf_sal  = soda_ssal,
  surf_temp = oisst_monthly)

# Now mask them all and get timeseries
masked_vars <- imap(observed_vars, function(masking_var, var_name){
  
  # Mask and get timeseries
  masked_gom   <- mask_shape(masking_var, trawl_gom)
  masked_gom   <- timeseries_from_mask(masked_gom, var_name)
  masked_trawl <- mask_shape(masking_var, trawl_full)
  masked_trawl <- timeseries_from_mask(masked_trawl,var_name)
  masked_dfo   <- mask_shape(masking_var, dfo_area)
  masked_dfo   <- timeseries_from_mask(masked_dfo, var_name)
    
  # Put in list
  list(
    nefsc_gom  = masked_gom,
    nefsc_full = masked_trawl,
    dfo_full   = masked_dfo)
  
})


# Get Time Periods to SODA/OISST
soda_years  <- unique(str_sub(names(soda_ssal), 2, 5))
oisst_years <- unique(str_sub(names(oisst_monthly), 2, 5))


# clean up
rm(soda_bsal, soda_btemp, soda_ssal, oisst_monthly)
```





### CMIP6 Bias-Corrected {.tabset}

#### Mean
```{r means}

# CMIP Bias Corrected
cmip_stemp <- stack(str_c(cmip_path, "BiasCorrected/surf_temp_OISST_bias_corrected_mean.grd"))
cmip_btemp <- stack(str_c(cmip_path, "BiasCorrected/bot_temp_SODA_bias_corrected_mean.grd"))
cmip_ssal  <- stack(str_c(cmip_path, "BiasCorrected/surf_sal_SODA_bias_corrected_mean.grd"))
cmip_bsal  <- stack(str_c(cmip_path, "BiasCorrected/bot_sal_SODA_bias_corrected_mean.grd"))


# Cmip matches to SODA/OISST Time Periods
soda_index  <- which(str_sub(names(cmip_bsal), 2, 5) %in% soda_years)
oisst_index <- which(str_sub(names(cmip_stemp), 2, 5) %in% oisst_years)


# Put them in a list
cmip_vars <- list(
  bot_temp  = cmip_btemp,
  bot_sal   = cmip_bsal,
  surf_sal  = cmip_ssal,
  surf_temp = cmip_stemp)

# Rotate them:
cmip_vars <- map(cmip_vars, raster::rotate)



# Now mask them all and get timeseries
masked_cmip <- imap(cmip_vars, function(masking_var, var_name){
  
  # Mask and get timeseries
  masked_gom   <- mask_shape(masking_var, trawl_gom)
  masked_gom   <- timeseries_from_mask(masked_gom, var_name)
  masked_trawl <- mask_shape(masking_var, trawl_full)
  masked_trawl <- timeseries_from_mask(masked_trawl,var_name)
  masked_dfo   <- mask_shape(masking_var, dfo_area)
  masked_dfo   <- timeseries_from_mask(masked_dfo, var_name)
    
  # put in list
  list(
    nefsc_gom  = masked_gom,
    nefsc_full = masked_trawl,
    dfo_full   = masked_dfo)
  
})
```

#### 5th Percentile

```{r}
# CMIP Bias Corrected
cmip_stemp_5 <- stack(str_c(cmip_path, "BiasCorrected/surf_temp_OISST_bias_corrected_5thpercentile.grd"))
cmip_btemp_5 <- stack(str_c(cmip_path, "BiasCorrected/bot_temp_SODA_bias_corrected_5thpercentile.grd"))
cmip_ssal_5  <- stack(str_c(cmip_path, "BiasCorrected/surf_sal_SODA_bias_corrected_5thpercentile.grd"))
cmip_bsal_5  <- stack(str_c(cmip_path, "BiasCorrected/bot_sal_SODA_bias_corrected_5thpercentile.grd"))


# Put them in a list
cmip_vars_5 <- list(
  bot_temp  = cmip_btemp_5,#[[soda_index]],
  bot_sal   = cmip_bsal_5,#[[soda_index]],
  surf_sal  = cmip_ssal_5,#[[soda_index]],
  surf_temp = cmip_stemp_5#[[oisst_index]]
  )

# Rotate them:
cmip_vars_5 <- map(cmip_vars_5, raster::rotate)

# Now mask them all and get timeseries
masked_cmip_5 <- imap(cmip_vars_5, function(masking_var, var_name){
  
  # Mask and get timeseries
  masked_gom   <- mask_shape(masking_var, trawl_gom)
  masked_gom   <- timeseries_from_mask(masked_gom, var_name)
  masked_trawl <- mask_shape(masking_var, trawl_full)
  masked_trawl <- timeseries_from_mask(masked_trawl,var_name)
  masked_dfo   <- mask_shape(masking_var, dfo_area)
  masked_dfo   <- timeseries_from_mask(masked_dfo, var_name)
    
  # put in list
  list(
    nefsc_gom  = masked_gom,
    nefsc_full = masked_trawl,
    dfo_full   = masked_dfo
  )
  
})
```


#### 95th Percentile

```{r}
# CMIP Bias Corrected
cmip_stemp_95 <- stack(str_c(cmip_path, "BiasCorrected/surf_temp_OISST_bias_corrected_95thpercentile.grd"))
cmip_btemp_95 <- stack(str_c(cmip_path, "BiasCorrected/bot_temp_SODA_bias_corrected_95thpercentile.grd"))
cmip_ssal_95  <- stack(str_c(cmip_path, "BiasCorrected/surf_sal_SODA_bias_corrected_95thpercentile.grd"))
cmip_bsal_95  <- stack(str_c(cmip_path, "BiasCorrected/bot_sal_SODA_bias_corrected_95thpercentile.grd"))


# Put them in a list
cmip_vars_95 <- list(
  bot_temp  = cmip_btemp_95,#[[soda_index]],
  bot_sal   = cmip_bsal_95,#[[soda_index]],
  surf_sal  = cmip_ssal_95,#[[soda_index]],
  surf_temp = cmip_stemp_95#[[oisst_index]]
  )

# Rotate them:
cmip_vars_95 <- map(cmip_vars_95, raster::rotate)


# Now mask them all and get timeseries
masked_cmip_95 <- imap(cmip_vars_95, function(masking_var, var_name){
  
  # Mask and get timeseries
  masked_gom   <- mask_shape(masking_var, trawl_gom)
  masked_gom   <- timeseries_from_mask(masked_gom, var_name)
  masked_trawl <- mask_shape(masking_var, trawl_full)
  masked_trawl <- timeseries_from_mask(masked_trawl,var_name)
  masked_dfo   <- mask_shape(masking_var, dfo_area)
  masked_dfo   <- timeseries_from_mask(masked_dfo, var_name)
    
  # put in list
  list(
    nefsc_gom  = masked_gom,
    nefsc_full = masked_trawl,
    dfo_full   = masked_dfo
  )
  
})


```


# Validating Seasonality {.tabset .tabset-pills}

The goal of the bias correction methods is to (as the name implies) remove bias from the CMIP6 model runs, through comparison to real-world observations or reanalysis data sets that better reflect real-world data. 

The bias-corrected data should mirror the intra-annual variations seen in the datasets used to bias-correct them, and the values should be close, and without an obvious bias.

The following plots showcase how the observational datasets fall in-line with the bias corrected means, 5th percentile, and 95th percentile data. With emphasis on intra-annual variations.

```{r}
comparison_plot <- function(var_option, mask_option){
  
  obs_data  <- masked_vars[[var_option]][[mask_option]]
  cmip_mean <- masked_cmip[[var_option]][[mask_option]]
  cmip_5    <- masked_cmip_5[[var_option]][[mask_option]]
  cmip_95   <- masked_cmip_95[[var_option]][[mask_option]]
  
  # Add data source
  cmip_mean$data_source <- "CMIP Mean"
  cmip_5$data_source    <- "CMIP 5th Perc."
  cmip_95$data_source   <- "CMIP 95th Perc."
  
  
    # Separate action for labeling oisst/soda
  if(var_option == "surf_temp"){
    obs_data$data_source <- "OISSTv2"

    } else{
   obs_data$data_source <- "SODA"
    }
  
  
  # Filter Dates to focus on shared timeperiod
  cmip_mean <- filter(cmip_mean, between(date, min(obs_data$date), max(obs_data$date)))
  cmip_5 <- filter(cmip_5, between(date, min(obs_data$date), max(obs_data$date)))
  cmip_95 <- filter(cmip_95, between(date, min(obs_data$date), max(obs_data$date)))
  
  
  # format variable for display on axis
  var_label <- str_replace(var_option, "_", " ")
  var_label <- str_to_title(var_label)
  
  ggplot() +
     geom_line(data = obs_data, aes_string("date", var_option, color = "data_source")) +
     geom_line(data = cmip_mean, aes_string("date", var_option, color = "data_source")) +
     geom_line(data = cmip_5, aes_string("date", var_option, color = "data_source")) +
     geom_line(data = cmip_95, aes_string("date", var_option, color = "data_source")) +
     scale_color_gmri() +
     labs(x = "", y = var_label, color = "") +
     facet_zoom(xlim =c(as.Date("2000-01-01"), as.Date("2001-12-31")), zoom.size = .5) +
    theme(legend.position = "bottom")
  
}
```


## Bottom Temperature {.tabset}

```{r}
var_option <- "bot_temp"


```


### Gulf of Maine

```{r}
mask_option <- "nefsc_gom"

comparison_plot(var_option = var_option, mask_option = mask_option)
```


### NMFS Survey Area

```{r}
mask_option <- "nefsc_full"

comparison_plot(var_option = var_option, mask_option = mask_option)
```

### DFO Survey Area

```{r}
mask_option <- "dfo_full"

comparison_plot(var_option = var_option, mask_option = mask_option)
```


## Surface Temperature {.tabset}

```{r}
var_option = "surf_temp"
```


### Gulf of Maine

```{r}
mask_option <- "nefsc_gom"

comparison_plot(var_option = var_option, mask_option = mask_option)
```


### NMFS Survey Area

```{r}
mask_option <- "nefsc_full"

comparison_plot(var_option = var_option, mask_option = mask_option)
```

### DFO Survey Area

```{r}
mask_option <- "dfo_full"

comparison_plot(var_option = var_option, mask_option = mask_option)
```



## Bottom Salinity {.tabset}

```{r}
var_option <- "bot_sal"

```


### Gulf of Maine

```{r}
mask_option <- "nefsc_gom"

comparison_plot(var_option = var_option, mask_option = mask_option)
```


### NMFS Survey Area

```{r}
mask_option <- "nefsc_full"

comparison_plot(var_option = var_option, mask_option = mask_option)
```

### DFO Survey Area

```{r}
mask_option <- "dfo_full"

comparison_plot(var_option = var_option, mask_option = mask_option)
```


## Surface Salinity {.tabset}

```{r}
var_option <- "surf_sal"


```


### Gulf of Maine

```{r}
mask_option <- "nefsc_gom"

comparison_plot(var_option = var_option, mask_option = mask_option)
```


### NMFS Survey Area

```{r}
mask_option <- "nefsc_full"

comparison_plot(var_option = var_option, mask_option = mask_option)
```

### DFO Survey Area

```{r}
mask_option <- "dfo_full"

comparison_plot(var_option = var_option, mask_option = mask_option)
```


# Regional Trajectories {.tabset .tabset-pills}

The following plots showcase how the observational datasets fall in-line with the bias corrected means, 5th percentile, and 95th percentile data sets.

```{r}

# Function to Plot Single Variable Trajectories for the Different Regions
clim_futures <- function(var_option){
  
  # Subset the variables
  obs_data  <- masked_vars[[var_option]]
  cmip_mean <- masked_cmip[[var_option]]
  cmip_5    <- masked_cmip_5[[var_option]]
  cmip_95   <- masked_cmip_95[[var_option]]
  
  # Label the sources for the CMIP Data, process areas
  
  obs_data  <- map_dfr(obs_data, ~ mutate(.x, data_source = "Reference Data"), .id = "area")
  cmip_mean <- map_dfr(cmip_mean, ~ mutate(.x, data_source = "CMIP Mean"), .id = "area")
  cmip_5    <- map_dfr(cmip_5, ~ mutate(.x, data_source = "CMIP 5th Perc."), .id = "area")
  cmip_95   <- map_dfr(cmip_95, ~ mutate(.x, data_source = "CMIP 95th Perc."), .id = "area")
  
  # combine the different quantiles
  cmip_all <- bind_rows(list(obs_data, cmip_mean, cmip_5, cmip_95)) %>% 
    mutate(data_source = factor(data_source, 
                                levels = c("Reference Data", "CMIP 5th Perc.", 
                                           "CMIP Mean", "CMIP 95th Perc.")),
           area = str_replace_all(area, "_", " "),
           area = str_to_title(area),
           area = str_replace(area, "Dfo", "DFO"),
           area = str_replace(area, "Nefsc", "NEFSC"),
           area = factor(area, levels = c("DFO Full", "NEFSC Gom", "NEFSC Full")),
           yr = lubridate::year(date),
           month = lubridate::month(date)) 
  
  
  #group on year and get annual averayges
  var_sym <- sym(var_option)
  cmip_all <- cmip_all %>% 
    group_by(yr, area, data_source) %>% 
    summarise({{ var_sym }} := mean({{ var_sym }}, na.rm = T),
              .groups = "keep") %>% 
    ungroup() %>% 
    mutate(yr = as.numeric(as.character(yr))) %>% 
    filter(yr >= 1982)
  
  
  # format variable for display on axis
  var_titles <- c(
    "bot_temp" = expression("Bottom Temperature"~degree~"C"),
    "surf_temp" = expression("Surface Temperature"~degree~"C"),
    "bot_sal" = "Bottom Salinity",
    "surf_sal" = "Surface Salinity")
  var_label <- var_titles[var_option]
  
  
  
  #pivot the sources wider so we can single them out asier
  data_wide <- cmip_all %>% 
    pivot_wider(names_from = data_source, values_from = {{var_option}})
  
  
  # Build plot
  regional_comparison_plot <-  ggplot(data_wide, aes(x = yr)) +
      geom_ribbon(aes(ymin = `CMIP 5th Perc.`, 
                      ymax = `CMIP 95th Perc.`),
                  fill = "gray80") +
      geom_line(aes(y = `CMIP Mean`, 
                    color = "CMIP Mean")) +
      geom_line(aes(y = `Reference Data`, color = "Reference Data")) +
      geom_point(aes(y = `Reference Data`, color = "Reference Data"),
                 size = 0.25) +
      scale_color_manual(values = c(
        "Reference Data" = "gray10", 
        "CMIP Mean"          = "gray50")) +
      facet_wrap(~area, nrow = 3) +
      labs(x = "", y = var_label, color = "") +
      theme(legend.position = "bottom")
  
  
  return(regional_comparison_plot)
    
  
  
}

```

## Bottom Temperature

```{r}
clim_futures(var_option = "bot_temp")
```


## Surface Temperature

```{r}
clim_futures(var_option = "surf_temp")
```


## Bottom Salinity

```{r}
clim_futures(var_option = "bot_sal")
```



## Surface Salinity

```{r}
clim_futures(var_option = "surf_sal")
```
