---
title: "Bias Correction Validations"
author: "Adam A. Kemberling"
date: "3/26/2021"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: hide
editor_options: 
  chunk_output_type: console

---

```{r setup, include=FALSE}
# Set knitr options
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, comment = NA)
options(knitr.kable.NA = '')

###__ Packages  ####
library(here)
library(ncdf4)
library(RNetCDF)
library(raster)
library(janitor)
library(gmRi)
library(patchwork)
library(tidyverse)
library(knitr)
library(sf)



# Load the build code and stratification function
box_paths  <- research_access_paths(os.use = "unix")
res_path   <- box_paths$res
oisst_path <- box_paths$oisst_mainstays
cmip_path  <- shared.path("unix", "RES_Data", "CMIP6")
soda_path  <- shared.path("unix", "RES_Data", "SODA")

#### Set theme  ####
theme_set(theme_minimal())

#  color palette for quick raster displays
temp_pal <- rev(RColorBrewer::brewer.pal(n = 10, name = "RdBu"))

####  Functions  ####
source(here("CMIP6_processing/R/sdm_workflow_funs.R"))
```



`r use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")`


# Validating Bias-Correction of CMIP6 Data

Bias corrected timelines for the Northeast Shelf will be compared against the reference observations (OISSTv2, SODA) to check how accurrately they represent observed conditions.


## Load Data

**SODA**

```{r}

# gom mask
gom_mask <- read_sf(str_c(res_path, "Shapefiles/NELME_regions/GoM_sf.shp"))


# SODA vars
soda_ssal  <- stack(str_c(soda_path, "SODA_Salt_Red.nc"))
soda_bsal  <- stack(str_c(soda_path, "SODA_Salt_Red_bottomLayer.nc"))
soda_btemp <- stack(str_c(soda_path, "SODA_Temp_Red_bottomLayer.nc"))

# masking function
mask_gom <- function(in_ras, gom_mask){
  r1 <- crop(in_ras, gom_mask)
  r2 <- mask(r1, gom_mask)
  return(r2)}

# mask them
soda_ssal  <- mask_gom(soda_ssal, gom_mask)
soda_bsal  <- mask_gom(soda_bsal, gom_mask)
soda_btemp <- mask_gom(soda_btemp, gom_mask)

# Plot Check
plot(soda_btemp[[1]], col = temp_pal)
```



**OISSTv2**

```{r}

# OISST 
data_window <- data.frame(lon = c(-72, -61),
                          lat = c(39, 46.1),
                          time = as.Date(c("1981-09-01", "2020-12-31")))

# Load Data
oisst <- oisst_window_load(oisst_path = oisst_path, 
                           data_window = data_window, 
                           anomalies = FALSE)


# Convert to Monthly
oisst <- stack(oisst)
month_key <- seq.Date(min(data_window$time), max(data_window$time), by = "month")
month_key <- str_c("X", month_key) %>% str_replace_all("-", ".")
month_key <- str_sub(month_key, 1, -4)
oisst_monthly <- map(month_key, function(wut_month){
  
  # Identify corresponding days in month
  which_layers  <- str_detect(names(oisst), wut_month)
  layer_indices <- which(which_layers)
  
  if(length(layer_indices) == 0) {
  print(str_c(length(layer_indices), " days for ", wut_month))
  }
  
  # Get mean
  month_mean <- calc(oisst[[layer_indices]], fun = mean, na.rm = T)

  # set month as name
  month_mean <- setNames(month_mean, wut_month)
  return(month_mean)
  
  
}) 

# Crop them
oisst_monthly <- stack(oisst_monthly)
oisst_monthly <- mask_gom(oisst_monthly, gom_mask = gom_mask)

# Plot check
plot(oisst_monthly[[1]], col = temp_pal)

```


**CMIP6 Bias-Corrected**

```{r}


# CMIP Bias Corrected
cmip_stemp <- stack(str_c(cmip_path, "BiasCorrected/surf_temp_OISST_bias_corrected_mean.grd"))
cmip_btemp <- stack(str_c(cmip_path, "BiasCorrected/bot_temp_SODA_bias_corrected_mean.grd"))
cmip_ssal  <- stack(str_c(cmip_path, "BiasCorrected/surf_sal_SODA_bias_corrected_mean.grd"))
cmip_bsal  <- stack(str_c(cmip_path, "BiasCorrected/bot_sal_SODA_bias_corrected_mean.grd"))

# Rotate them:
cmip_stemp <- raster::rotate(cmip_stemp)
cmip_btemp <- raster::rotate(cmip_btemp)
cmip_ssal  <- raster::rotate(cmip_ssal)
cmip_bsal  <- raster::rotate(cmip_bsal)


# Crop them all
cmip_stemp <- mask_gom(cmip_stemp , gom_mask)
cmip_btemp <- mask_gom(cmip_btemp , gom_mask)
cmip_ssal  <- mask_gom(cmip_ssal  , gom_mask)
cmip_bsal  <- mask_gom(cmip_bsal  , gom_mask)

# Plot Check
plot(cmip_stemp[[1]], col = temp_pal)

```
# Visual Comparisons {.tabset .tabset-pills}

**Processing and Plotting Functions**

```{r}
# Make dataframe using cellStats
ras_timeline <- function(named_ras){
  data.frame(ras_date = names(named_ras),
             mean_val = cellStats(named_ras, stat = "mean", na.rm = T)) %>% 
    mutate(ras_date = str_sub(ras_date, 2, -1),
           ras_date = str_replace_all(ras_date, "[.]", "/"),
           ras_date = str_c(ras_date, "/15"),
           ras_date = as.Date(ras_date))
}

# Plot Comparison
timeline_compare <- function(data1, data2, name1, name2, variable){ 
  ggplot() +
    geom_line(data = data1, aes(ras_date, mean_val, color = name1)) +
    geom_line(data = data2, aes(ras_date, mean_val, color = name2)) +
    scale_color_gmri() +
    labs(x = "", y = variable, color = "Data Source") +
    theme(legend.position = "bottom")
}
```



## Surface Temperature

```{r}

# Make timelines
cmip_stemp_tl <- ras_timeline(cmip_stemp)
oisst_tl      <- ras_timeline(oisst_monthly)

# Compare
timeline_compare(data1 = cmip_stemp_tl, 
                 data2 = oisst_tl, 
                 name1 = "CMIP Bias Corrected", 
                 name2 = "OISSTv2", 
                 variable = "Surface Temperature")


```


## Surface Salinity

```{r}
# Make timelines
cmip_ssal_tl <- ras_timeline(cmip_ssal)
soda_ssal_tl <- ras_timeline(soda_ssal)

# Compare
timeline_compare(data1 = cmip_ssal_tl, 
                 data2 = soda_ssal_tl, 
                 name1 = "CMIP Bias Corrected", 
                 name2 = "SODA", 
                 variable = "Surface Salinity")
```


## Bottom Temperature

```{r}
# Make timelines
cmip_btemp_tl <- ras_timeline(cmip_btemp)
soda_btemp_tl <- ras_timeline(soda_btemp)

# Compare
timeline_compare(data1 = cmip_btemp_tl, 
                 data2 = soda_btemp_tl, 
                 name1 = "CMIP Bias Corrected", 
                 name2 = "SODA", 
                 variable = "Bottom Temperature")
```


## Bottom Salinity

```{r}
# Make timelines
cmip_bsal_tl <- ras_timeline(cmip_bsal)
soda_bsal_tl <- ras_timeline(soda_bsal)

# Compare
timeline_compare(data1 = cmip_bsal_tl, 
                 data2 = soda_bsal_tl, 
                 name1 = "CMIP Bias Corrected", 
                 name2 = "SODA", 
                 variable = "Bottom Salinity")
```


`r insert_gmri_footer()`