---
title: "CMIP-OISST Bias Correciton"
author: "Adam A. Kemberling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: show
params: 
  experiment: 
      label: "Select CMIP experiment"
      value: "ssp1_26"
      input: select
      choices: ["ssp1_26", "ssp1_85"]
  climate_start: "1985"
  climate_stop: "2014"
editor_options: 
  chunk_output_type: console

---

```{r setup, include=FALSE}

# Set knitr options
knitr::opts_chunk$set(echo = TRUE, message = T, warning = F, comment = NA)
options(knitr.kable.NA = '')

###__ Packages  ####
library(here)
library(ncdf4)
library(RNetCDF)
library(raster)
library(janitor)
library(gmRi)
library(patchwork)
library(tidyverse)
library(knitr)



# Load the build code and stratification function
box_paths  <- research_access_paths() # os.use = "unix" Doesn't seem to be an argument??
res_path   <- box_paths$res
oisst_path <- box_paths$oisst_mainstays
cmip_path  <- shared.path("unix", "RES_Data", "CMIP6")

#### Set theme  ####
theme_set(theme_minimal())

#  color palette for quick raster displays
temp_pal <- rev(RColorBrewer::brewer.pal(n = 10, name = "RdBu"))

####  Functions  ####
source(here("CMIP6_processing/R/sdm_workflow_funs.R"))

experiment <- params$experiment
cmip_date_key <- cmip_date_key_fun(experiment)
```



`r use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")`


## Load CMIP6 Data for sea surface Temperature

 
The only variable we will use OISSTv2 data to bias correct for is sea surface temperature.



```{r load collection}

####  Choose a variable
pick_var <- "surf_temp"


# Load all the CMIP Scenarios - cropped to study area
cmip_data <- import_cmip_collection(cmip_var = pick_var, experiment)
```


### Match Dates to Xarray

Sometimes the raster package has an issue with the dates used by the climate models, more specifically how its a non-standard calendar without leap years. Xarray doesn't really have an issue for some reason, so this step matches the file to what the dates should be, to ensure proper date comparisons down the line.

```{r get layer names}

# Get the cftime dates from the date key - this ensures proper date matching
# the fallback is to assign year and month generally based on if its an historic run or not
cmip_data <- imap(.x = cmip_data, 
                  .f = function(cmip_dataset, cmip_source){
  message(paste("Setting Names for: ", cmip_source))
  names(cmip_dataset) <- get_cmip_dates(cmip_source = cmip_source, 
                                        cmip_var = pick_var,
                                        time_dim = dim(cmip_dataset)[3])
  return(cmip_dataset)})

```



```{r}
plot(cmip_data[[1]][[1]],
     main = "Single Month of CMIP6 Data",
     col = temp_pal)
```

### Screening Strange Values

There is the possibility that in some data sources or different raster layers within them that values that are exceptionally high or low may be used in place of an explicit `NA`. These will be issues when we look across models to get summary metrics.


```{r}
# Checking min and max values

# # can use clamp to set limits around plausible values
# clamp_test <- clamp(cmip_data[[1]], lower = -50, upper = 100, useValues = FALSE)

# boiling water seems like a reasonable limit
cmip_data <- imap(cmip_data, function(ras_brick, ras_lab){ 
  message(paste0("Screening Values for:  ", ras_lab))
  clamp(ras_brick, lower = -50, upper = 100, useValues = FALSE) 
})

```







### Build Lookup Table of Historic and Projected Climate Runs

The CMIP to anomalies function relies on a consistent naming structure of XYYYYMMDD to pull out the correct years and months. If this isn't the case the function as currently written will break.

This table is also used to match climatologies to their historic periods.

```{r name structure table}
#####  Build table of name structure and length dimension and file source type


#### 1. Get the start and end time dimensions that R gives
tdims <- map_dfr(cmip_data, function(cmip_stack){
    # Pull out some descriptive aspects of each
    # Assemble as a table
    time_dims  <- dim(cmip_stack)[3]
    summary_table <- data.frame( "time_dim"  = time_dims)}
    # collapse to dataframe
    , .id = "file_name")


#### 2. Do some formatting and rearranging to label historic etc.
tdims <- tdims %>%
  mutate(
    scenario_type = ifelse(str_detect(file_name, "historic"), "historic", "projection"),
    cmip_scenario = ifelse(str_detect(file_name, "historic"),
                           str_remove(file_name, "_historical"),
                           str_remove(file_name, "_ssp...")),
    nc_name = paste0(file_name, ".nc")) %>% 
  select(cmip_scenario, scenario_type, time_dim, file_name, nc_name)

# The code digging into the dates is no longer relevant, just need the scenario type info
name_structure <- tdims
```


## Build CMIP Climatologies

Need two things now:

 1. CMIP6 Climatology for the variables we loaded
 2. The corresponding anomalies


```{r}

# Set start and end year for climatology - needs to match SODA climatology reference period
start_year <- params$climate_start
end_year   <- params$climate_stop


# Determine which netcdf files cover the reference period and should be used for the climatologies
historic_runs <- which(str_detect(names(cmip_data), "historic"))


# Run the climatologies for the historic runs only
cmip_clims <- imap(cmip_data[historic_runs], function(cmip_stack, cmip_name){
  
  # Get monthly climatology
  message(paste0("Processing Climatology: ", cmip_name))
  cmip_clim <- suppressMessages(cmip_to_clim(cmip_stack = cmip_stack, 
                                             clim_years = c(start_year, end_year)))
  
  # return the climatology (should work b/c smaller)
  return(cmip_clim)})


```



```{r}
plot(cmip_clims[[1]][[1]],
     main = "Single Month of a CMIP6 Climatology",
     col = temp_pal)
```



## Get CMIP Scenario Anomalies (deltas)

Now the climatologies are matched up with the correct data sets to get anomalies for both the historic periods as well as the matching climate projections.


```{r get cmip anomalies}
cmip_anoms <- imap(cmip_data, function(cmip_data, cmip_name){
  
  # Check for those trouble files
  if(class(cmip_clims) == "character"){
    return("Problem with CMIP Naming Structure")}
  
  
  # Identify root cmip file & its number in the list
  cmip_root <- str_replace(cmip_name, "_ssp...", "_historical")
  clim_id <- which(names(cmip_clims) == cmip_root)
  
  # # Get the matching climatology
  if(length(clim_id) != 0){
    
    # Grab the climatology
    clim_use <- cmip_clims[[clim_id]]

    # print progress
    message(paste0(cmip_name, " has matching climatology, processing anomalies."))
    
    # Use sdm_workflow function to match months and return anomalies
    cmip_anomalies <- cmip_get_anomalies(cmip_data, clim_use)
    return(cmip_anomalies)
    
  } else if(length(clim_id) == 0) {
    message(paste0("No matching climate reference for: ", cmip_name))
    cmip_anomalies <- "ugly duckling"
    return(cmip_anomalies)}
  
})




# Pull out the cases without historical matches
no_matches <- which(map_chr(cmip_anoms, class) == "character")
good_data <- which(names(cmip_anoms) %not in% names(no_matches))

# and subset
cmip_anoms <- cmip_anoms[good_data]


# clean up memory
rm(cmip_data)
```


```{r}
# Plot Check
plot(cmip_anoms[[1]][[1]], 
     main = "Single Month of CMIP6 Anomalies",
     col = temp_pal)
```


##  OISSTv2 Bias Correction

In order to do this we need the climatology for each of those variables for a reference period that aligns with the CMIP historical data


### Import OISST Climatology

The OISST climatology was processed using xarray in the oisst mainstays repository. Once it has been loaded we also need to convert it to monthly to match the CMIP data.

```{r import ref climatology}

# Load OISST Climatology
oisst_clim <- import_oisst_clim(climatology_period = c("1985-2014"), 
                                os.use = "unix")



# # Use clamp if necessary?
# range_mets <- range(oisst_clim)
# range(values(range_mets$range_min), na.rm = T) # how annoying it is to just get 2 values
# range(values(range_mets$range_max), na.rm = T) # how annoying it is to just get 2 values
```

### Convert to Monthly Means

```{r}
# Convert to Monthly
oisst_monthly <- months_from_clim(oisst_clim)
rm(oisst_clim)
```

### Optional re-gridding

This step  is only necessary if the resolution across data sources is not the same. The re-sample grid function just takes an input gridded dataset and re-samples it to match the resolution of a second gridded dataset, uses bi-linear interpolation for re-sampling.

```{r}
# ####  Check/Resample Grids  ####
# not needed since Matt does this ahead of time


# cmip_anoms <- map(cmip_anoms, function(anom_grid){
#   if(class(anom_grid) == "character"){return("Problem with CMIP Naming Structure")}
#   resample_grid(starting_grid = anom_grid, desired_grid = oisst_monthly)})

```

```{r}
# Plot Check
plot(oisst_monthly[[1]], 
     col = temp_pal, 
     main = "January Climatology")
```



### Bias Correction

This is the step where the anomalies/deltas from each CMIP6 run are applied to the climatology of the reference data climatology, either SODA/OISST.

```{r perform bias correction}
####  Match and Bias Correct
cmip_anoms_bcorrect <- map(cmip_anoms, function(anom_grid){
  
  # check for problem data
  if(class(anom_grid) == "character"){
    return("Problem with CMIP Naming Structure")}
  
  # run for data that passes check
  delta_method_bias_correct(cmip_grid = anom_grid, 
                            reference_climatology = oisst_monthly)})




# remove regular anomalies
rm(cmip_anoms)
```




```{r}
# Single Run
plot(cmip_anoms_bcorrect[[1]][[1]], 
     main = "Single Month of Bias Corrected Data", 
     col = temp_pal)
```



### Saving Bias Corrected CMIP6 Runs

```{r, eval = FALSE}
# iwalk(cmip_anoms_bcorrect, function(bias_corrections, run_name){
#    paste0(cmip_path, "BiasCorrected/", pick_var)
#    out_location <- paste0(cmip_path, "BiasCorrected/", pick_var, "/", run_name, ".grd")
#    message(paste0("Saving ", run_name, " at: \n", out_location))
#    # writeRaster(bias_corrections, out_location, overwrite = T)
#   
# })
```





## Processing Mean/5th/95th

The following code sets up the sub-setting and matching components to process comparable model runs. Since the historical period and future projections cover different periods, they get run separately here, to be combined again after different data representing across-model quantiles have been pulled from them.

```{r quantile setup}

####  Different Treatments for Historical / Projections

# Get the names for historic and future runs
historic_names     <- filter(name_structure, scenario_type == "historic") %>% pull(file_name)
projection_names   <- filter(name_structure, scenario_type == "projection") %>% pull(file_name)

# Their matching index numbers
historic_sources   <- which(names(cmip_anoms_bcorrect) %in% historic_names)
projection_sources <- which(names(cmip_anoms_bcorrect) %in% projection_names)

# Use those to separate the two groups
historic_bias_corr   <- cmip_anoms_bcorrect[historic_sources]
projection_bias_corr <- cmip_anoms_bcorrect[projection_sources]

# Will need a year-month key for all the time steps for historic and projected data
historic_key   <- names(historic_bias_corr[[1]]) %>% str_sub(1,8)
projection_key <- names(projection_bias_corr[[1]]) %>% str_sub(1,8)

# can also manufacture them since we have a thousand time checked the dates



```


### Process Historical Period Climate Quantiles

Now that all the CMIP6 data has been bias corrected using either OISST or SODA, we can take the bias corrected data and build arrays that capture the variability across all models by pulling out values at each time step that represent different measures of the collection. So at each cell for every time step we now take the value from whichever model represents the 5th percentile, the mean, or the 95th percentile.

```{r historic quants}

####  Historic Quantiles


# Use the time-period length (in months) to Get the mean 5th 95th at each step
historic_mean <- time_period_quantile(time_period = "historic",
                                      time_period_collection = historic_bias_corr, 
                                      quantile_product = "mean")



# 5th Percentile
historic_05 <- time_period_quantile(time_period = "historic",
                                    time_period_collection = historic_bias_corr, 
                                    quantile_product = "5th")


# 95th Percentile
historic_95 <- time_period_quantile(time_period = "historic",
                                    time_period_collection = historic_bias_corr, 
                                    quantile_product = "95th")



# clean up
rm(historic_bias_corr)
```


```{r, fig.height = 8}
# Plot check
par(mfrow = c(3, 1))
plot(historic_mean$X1950.01, main = "Historic Mean", col = temp_pal)
plot(historic_05$X1950.01,   main = "Historic 5th Percentile", col = temp_pal)
plot(historic_95$X1950.01,   main = "Historic 95th Percentile", col = temp_pal)
```


### Process Climate Projection Period Quantiles

Repeat what we just did for the historical periods for the future projections. At each cell for every time step get the value from whichever model represents the 5th percentile, the mean, or the 95th percentile.

```{r projection quants}

####  Projection Quantiles


# Mean
projection_mean <- time_period_quantile(time_period = "projection",
                                    time_period_collection = projection_bias_corr, 
                                    quantile_product = "mean")


# 5th Percentile
projection_05 <- time_period_quantile(time_period = "projection",
                                     time_period_collection = projection_bias_corr, 
                                     quantile_product = "5th")


# 9th Percentile
projection_95 <- time_period_quantile(time_period = "projection",
                                    time_period_collection = projection_bias_corr, 
                                    quantile_product = "95th")


# clean up
rm(projection_bias_corr)
```


```{r, fig.height=8}
# Plot check
par(mfrow = c(3,1))
plot(projection_mean$X2015.01, main = "Projection Mean", col = temp_pal)
plot(projection_05$X2015.01,   main = "Projection 5th Percentile", col = temp_pal)
plot(projection_95$X2015.01,   main = "Projection 95th Percentile", col = temp_pal)

```

### Assemble complete Timelines

Once the different quantiles have been run for both the historical periods and the future projections, these two periods can be joined together to create a complete stack with time from 1950-2100.

```{r}
full_5th  <- stack(stack(historic_05), stack(projection_05))
full_mean <- stack(stack(historic_mean), stack(projection_mean))
full_95th <- stack(stack(historic_95), stack(projection_95))
```


## Export Quantiles

```{r}
# Path to the bias correction folder
var_path <- paste0(cmip_path, experiment, "/BiasCorrected/EnsembleData/", pick_var,"/", pick_var)


# File Names
writeRaster(full_5th, paste0(var_path, "_OISST_bias_corrected_5thpercentile.grd"), overwrite = T)
writeRaster(full_mean, paste0(var_path, "_OISST_bias_corrected_mean.grd"), overwrite = T)
writeRaster(full_95th, paste0(var_path, "_OISST_bias_corrected_95thpercentile.grd"), overwrite = T)
```


`r insert_gmri_footer()`